Multi-product Batch Processing Time Maximization Problem

### 2023-08-14 manuscriptBlinded.tex
paper before submiting

%\author{Tatiana Balbi Fraga, Ítalo Ruan Barbosa de Aquino and Regilda da Costa e Silva Menêzes}

\author{Tatiana Balbi Fraga\corref{cor1}\fnref{label1}}
\ead{tatiana.balbi@ufpe.br}
\cortext[cor1]{corresponding author}

\author{Ítalo Ruan Barbosa de Aquino\fnref{label1}}
\ead{italo_ruan_@hotmail.com}

\author{Regilda da Costa e Silva Menêzes\fnref{label1}}
\ead{regilda.smenezes@ufpe.br}

@@

\affiliation[label1]{
		   organization={Agreste Academic Center - Federal University of Pernambuco},%Department and Organization
           addressline={Avenida Marielle Franco, Nova Caruaru}, 
           city={Caruaru},
           postcode={55014-900}, 
           state={PE},
           country={Brazil}}

@@

Starting a working day... 
Tasks finish tests using LINGO on optimization mode.
Submit paper again to journal.

@@

\begin{algorithm}
\caption{Solving MBPTMP | Part 01 - find an optimal $T$ to the problem defined in (\ref{MBPTMP01}) to (\ref{MBPTMP10}).}\label{alg:part01}
\begin{algorithmic}
	\Require $\textrm{UD}_i, \textrm{UO}_i, \textrm{UI}_i, \textrm{p}_i, \forall i, \textrm{O}, \textrm{I}, \textrm{Z}$
	\State $T^* \gets \lfloor{\min \{\min_{\forall i} \{(\textrm{UO}_i + \textrm{UI}_i + \textrm{UD}_i) / \textrm{p}_i\},(\textrm{O} + \textrm{I} + \sum_i \textrm{UD}_i) / \sum_i {\textrm{p}_i}\}}\rfloor$
	\State $T \gets \min \{T^* , \textrm{Z}\}$
\State \textbf{Return:} $T$.
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{Solving MBPTMP | Part 02 - calculate $D_i$, $O_i$ and $I_i$, $\forall i$, ignoring the restrictions in (\ref{eq:resO}) and (\ref{eq:resI}).}\label{alg:part02}
\begin{algorithmic}
	\Require $\textrm{UD}_i, \textrm{UO}_i, \textrm{UI}_i, \textrm{p}_i, \forall i, \textrm{O}, \textrm{I}, \textrm{Z}, T$.
	\For{all i}
		\State $E_i \gets T * \textrm{p}_i$
		\State $D_i \gets \min \{\textrm{UD}_i, E_i\}$
		\State $E_i \gets E_i - D_i$
		\State $O_i \gets \min \{\textrm{UO}_i, E_i\}$
		\State $E_i = E_i - O_i$
		\State $\textrm{UO}_i = \textrm{UO}_i - O_i$
		\State $I_i = E_i$
		\State $\textrm{UI}_i = \textrm{UI}_i - I_i$
	\EndFor
	\State $SO = \textrm{O} - \sum_i{O_i}$
	\State $SI = \textrm{I} - \sum_i{I_i}$
	\State \textbf{Return:} $D_i, O_i, I_i, E_i, \textrm{UO}_i, \textrm{UI}_i, \forall i, SO, SI$.
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{Solving MBPTMP | Part 03: redistribute production to comply with limitation restrictions for the batch products set.}\label{alg:part03}
\begin{algorithmic}
\Require $O_i, I_i, \textrm{UO}_i, \textrm{UI}_i, \forall i, SO, SI$.
\If{$SO < 0$}
\For{all $i$}
	\State $O_i \gets O_i - \min\{ O_i, UI_i, |SO|\}$
	\State $I_i \gets I_i + \min\{ O_i, UI_i, |SO|\}$
	\State $SO \gets SO + \min\{ O_i, UI_i, |SO|\}$
	\If{$SO = 0$}
		break looping for;
	\EndIf
\EndFor
\EndIf

@@

\emph{Solution for the example presented before - determining the values of $D_i$, $O_i$ and $I_i$, $\forall i$.}

$T = 55 \textrm{min}$

$E_A = 55 \textrm{min}* 60 \textrm{ g}/\textrm{min}  = 3,300 \textrm{ g}$

$E_B = 55 \textrm{min}* 40 \textrm{ g}/\textrm{min}  = 2,200 \textrm{ g}$

$D_A = \min \{1,000 \textrm{ g}, \textrm{ } 3,300 \textrm{ g}\} = 1,000 \textrm{ g} $

$D_B = \min \{500 \textrm{ g}, \textrm{ } 2,200 \textrm{ g}\} = 500 \textrm{ g} $

$E_A = 3,300 \textrm{ g} - 1,000 \textrm{ g} = 2,300 \textrm{ g}$

$E_B = 2,200 \textrm{ g} - 500 \textrm{ g} = 1,700 \textrm{ g}$

$O_A = \min \{600 \textrm{ g}, \textrm{ }  2,000 \textrm{ g}\} = 600 \textrm{ g}$

$O_B = \min \{600 \textrm{ g}, \textrm{ }  1,700 \textrm{ g}\} = 600 \textrm{ g}$

$E_A = 2,300 \textrm{ g} - 600 \textrm{ g} = 1,700 \textrm{ g}$

$E_B = 1,700 \textrm{ g} - 600 \textrm{ g} = 1,100 \textrm{ g}$

$\textrm{UO}_A = 600 \textrm{ g} - 600 \textrm{ g} = 0 \textrm{ g}$

$\textrm{UO}_B = 600 \textrm{ g} - 600 \textrm{ g} = 0 \textrm{ g}$

$I_A = 1,700 \textrm{ g}$

$I_B = 1,100 \textrm{ g}$

$\textrm{UI}_A = 3,000 \textrm{ g} - 1,700 \textrm{ g} = 1,300 \textrm{ g}$

$\textrm{UI}_B = 2,000 \textrm{ g} - 1,100 \textrm{ g} = 900 \textrm{ g}$

$SO = 1,000 \textrm{ g} - 1,200 \textrm{ g} = - 200 \textrm{ g}$

$SI = 3,000 \textrm{ g} - 2,800 \textrm{ g} = 200 \textrm{ g}$

\emph{Redistribution}

$O_A = 600 \textrm{ g} - \min \{600 \textrm{ g}, \textrm{ }  1,300 \textrm{ g}, \textrm{ }  200 \textrm{ g}\} = 400 \textrm{ g}$

$I_A = 1,700 \textrm{ g} + \min \{600 \textrm{ g}, \textrm{ }  1,300 \textrm{ g}, \textrm{ }  200 \textrm{ g}\} = 1,900 \textrm{ g}$

$SO = 200 \textrm{ g} + \min \{600 \textrm{ g}, \textrm{ }  1,300 \textrm{ g}, \textrm{ }  200 \textrm{ g}\} = 0 \textrm{ g}$

\emph{Solution}

$T = 55 \textrm{min}; \quad D_A = 1,000 \textrm{ g}; \quad D_B = 500 \textrm{ g}; \quad O_A = 400 \textrm{ g}; \quad O_B = 600 \textrm{ g}; \quad I_A = 1,900 \textrm{ g}; \quad \textrm{and } I_B = 1,100 \textrm{ g}$.

@@

To test the developed model and analytical solution method, we created a solver in C++ named COPSolver (V01\_20230809, lib multiproduct-batch-processing-time-maximization-problem) and a solver in LINGO named MPBPTM.lng both available at https://github.com/tbfraga/COPSolver. The tests were performed on a notebook with an Intel i7 processor. We tested the solvers developed for the benchmarks presented on Tables \ref{tab:MBPTMP001}, \ref{tab:MBPTMP002} and \ref{tab:MBPTMP003}, and for randon benchmarks generated by the following functions: \\

@@

Randomly generated benchmarks were named RMBPTM $\textrm{N}$, being $\textrm{N}$ the number of products. Seeking to enable the reproduction of the results, in the computational construction of the benchmarks we used the function srand((unsigned) source), where source is a defined value. To build the results presented in this work, we used source=0. The benchmarks used for the tests performed can be consulted at github.com/tbfraga/COPSolver.

@@

\begin{table}[h]
\begin{center}
\begin{footnotesize}
\begin{tabular}[c]{l r r r r}
problem & LINGO solver & time (s) & analytical method & time (s) \\
\cline {1-5} \\
MBPTM 2 & 55 & 0.04 & 55 & 0.001 \\
MBPTM 3 & 48 & 0.06 & 48 & 0.001 \\
MBPTM 10 & 30 & 0.09 & 30 & 0.001 \\
RMBPTM 20 & 100 & 0.10 & 100 & 0.002 \\
RMBPTM 50 & 98 & 0.07 & 98 & 0.002\\
RMBPTM 100 & 98 & 0.11 & 98 & 0.002\\
RMBPTM 1,000 & 78 & 9.59 & 78 & 0.013\\
RMBPTM 2,000 & 70 & 3.24 & 70 & 0.027\\
RMBPTM 5,000 & 70 & 15.28 & 70 & 0.079 \\
RMBPTM 10,000 & 70 & 58.22 & 70 & 0.185 \\
\cline {1-5} \\
\end{tabular}
\caption{Results obtained with the LINGO solver and the analytical method}
\label{tab:results}
\end{footnotesize}
\end{center}
\end{table}

Tables (\ref{tab:compResultsMBPTMP2}) and (\ref{tab:compResultsMBPTMP3}) shows the production distribution for MBPTM 3 and 10 found by both COPSolver and LINGO solver. Files containing all results can be found at github.com/tbfraga/COPSolver.

@@

T.B. Fraga: Conceptualization, Project administration, Supervision, Software, Methodology, Validation, Formal analysis, Writing – original draft, Writing – review \& editing. Í.R.B. Aquino: Data curation. R.C.S. Menêzes: Data curation.

>>>>

@@

@@

@@

\begin{algorithm}
\caption{Solving MBPTM problem | Part 01 - find an optimal $T$ to the problem defined in (\ref{MBPTMP01}) to (\ref{MBPTMP10}).}\label{alg:part01}
\begin{algorithmic}
	\Require $\textrm{UD}_i, \textrm{UO}_i, \textrm{UI}_i, \textrm{p}_i, \forall i, \textrm{O}, \textrm{I}, \textrm{Z}$
	\State $T^* \gets \lfloor{\min \{\min_{\forall i} \{(\textrm{UO}_i + \textrm{UI}_i + \textrm{UD}_i) / \textrm{p}_i\},(\textrm{O} + \textrm{I} + \sum_i \textrm{UD}_i) / \sum_i {\textrm{p}_i}\}}\rfloor$
	\State $T \gets \min \{T^* , \textrm{Z}\}$
\State \textbf{Return:} $T$.
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{Solving MBPTM problem | Part 02 - calculate $D_i$, $O_i$ and $I_i$, $\forall i$, ignoring the restrictions in (\ref{eq:resO}) and (\ref{eq:resI}).}\label{alg:part02}
\begin{algorithmic}
	\Require $\textrm{UD}_i, \textrm{UO}_i, \textrm{UI}_i, \textrm{p}_i, \forall i, \textrm{O}, \textrm{I}, \textrm{Z}, T$.
	\For{all i}
		\State $E_i \gets T * \textrm{p}_i$
		\State $D_i \gets \min \{\textrm{UD}_i, E_i\}$
		\State $E_i \gets E_i - D_i$
		\State $O_i \gets \min \{\textrm{UO}_i, E_i\}$
		\State $E_i = E_i - O_i$
		\State $\textrm{UO}_i = \textrm{UO}_i - O_i$
		\State $I_i = E_i$
		\State $\textrm{UI}_i = \textrm{UI}_i - I_i$
	\EndFor
	\State $SO = \textrm{O} - \sum_i{O_i}$
	\State $SI = \textrm{I} - \sum_i{I_i}$
	\State \textbf{Return:} $D_i, O_i, I_i, E_i, \textrm{UO}_i, \textrm{UI}_i, \forall i, SO, SI$.
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{Solving MBPTM problem | Part 03: redistribute production to comply with limitation restrictions for the batch products set.}\label{alg:part03}
\begin{algorithmic}
\Require $O_i, I_i, \textrm{UO}_i, \textrm{UI}_i, \forall i, SO, SI$.
\If{$SO < 0$}
\For{all $i$}
	\State $O_i \gets O_i - \min\{ O_i, UI_i, |SO|\}$
	\State $I_i \gets I_i + \min\{ O_i, UI_i, |SO|\}$
	\State $SO \gets SO + \min\{ O_i, UI_i, |SO|\}$
	\If{$SO = 0$}
		break looping for;
	\EndIf
\EndFor
\EndIf

@@

@@

To test the developed model and analytical solution method, we created a solver in C++ named COPSolver (V01\_20230814, lib multiproduct-batch-processing-time-maximization-problem) and a solver in LINGO named MBPTM.lng both available at https://github.com/Blinded. The tests were performed on a notebook with an Intel i7 processor. We tested the solvers developed for the benchmarks presented on Tables \ref{tab:MBPTMP001}, \ref{tab:MBPTMP002} and \ref{tab:MBPTMP003}, and for randon benchmarks generated by the following functions: \\

@@

Randomly generated benchmarks were named RMBPTM $\textrm{N}$, being $\textrm{N}$ the number of products. Seeking to enable the reproduction of the results, in the computational construction of the benchmarks we used the function srand((unsigned) source), where source is a defined value. To build the results presented in this work, we used source=0. The benchmarks used for the tests performed can be consulted at github.com/Blinded.

@@

\begin{table}[h]
\begin{center}
\begin{footnotesize}
\begin{tabular}[c]{l r r r r}
problem & LINGO solver & time (s) & analytical method & time (s) \\
\cline {1-5} \\
MBPTM 2 & 55 & 0.03 & 55 & $<$ 0.01 \\
MBPTM 3 & 48 & 0.03 & 48 & $<$ 0.01 \\
MBPTM 10 & 30 & 0.05 & 30 & $<$ 0.01 \\
RMBPTM 20 & 100 & 0.06 & 100 & $<$ 0.01 \\
RMBPTM 50 & 98 & 0.08 & 98 & $<$ 0.01 \\
RMBPTM 100 & 98 & 0.11 & 98 & $<$ 0.01 \\
RMBPTM 1,000 & 78 & 1.20 & 78 & $<$ 0.01 \\
RMBPTM 2,000 & 70 & 3.25 & 70 & $<$ 0.01 \\
RMBPTM 5,000 & 70 & 15.12 & 70 & $<$ 0.02 \\
RMBPTM 10,000 & 70 & 55.19 & 70 & $<$ 0.04 \\
\cline {1-5} \\
\end{tabular}
\caption{Results obtained with the LINGO solver and the analytical method}
\label{tab:results}
\end{footnotesize}
\end{center}
\end{table}

Tables (\ref{tab:compResultsMBPTMP2}) and (\ref{tab:compResultsMBPTMP3}) shows the production distribution for MBPTM 3 and 10 found by both COPSolver and LINGO solver. Files containing all results can be found at github.com/Blinded.

@@

Blinded
